### 2014-04-26 paper.tex
Add paper 3.
Up to around 1.25k words :-) possibly writing a bit too much for each paper,
but we'll see.

%% bare_conf.tex
%% V1.4
%% 2012/12/27
%% by Michael Shell
%% See:
%% http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.8 or later) with an IEEE conference paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/tex-archive/macros/latex/contrib/IEEEtran/
%% and
%% http://www.ieee.org/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE! 
%% User assumes all risk.
%% In no event shall IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%
%% File list of work: IEEEtran.cls, IEEEtran_HOWTO.pdf, bare_adv.tex,
%%                    bare_conf.tex, bare_jrnl.tex, bare_jrnl_compsoc.tex,
%%                    bare_jrnl_transmag.tex
%%*************************************************************************
\documentclass[conference]{IEEEtran}
% Add the compsoc option for Computer Society conferences.

\ifCLASSINFOpdf
  \usepackage[pdftex]{graphicx}
  \graphicspath{{../img/}}
  \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  \usepackage[dvips]{graphicx}
  \graphicspath{{../img/}}
  \DeclareGraphicsExtensions{.eps}
\fi

\usepackage[cmex10]{amsmath}

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/tools/

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

% *** SUBFIGURE PACKAGES ***
%\ifCLASSOPTIONcompsoc
%  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
%\else
%  \usepackage[caption=false,font=footnotesize]{subfig}
%\fi
% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
% for subfigure.sty, the latter of which is no longer maintained and is
% incompatible with some LaTeX packages including fixltx2e. However,
% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
% which will override IEEEtran.cls' handling of captions and this will result
% in non-IEEE style figure/table captions. To prevent this problem, be sure
% and invoke subfig.sty's "caption=false" package option (available since
% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
% handling of captions.
% Note that the Computer Society format requires a larger sans serif font
% than the serif footnote size font used in traditional IEEE formatting
% and thus the need to invoke different subfig.sty package options depending
% on whether compsoc mode has been enabled.
%
% The latest version and documentation of subfig.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/subfig/

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure. The latest version and documentation can be found at:
% http://www.ctan.org/tex-archive/macros/latex/base/

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/sttools/
% Do not use the stfloats baselinefloat ability as IEEE does not allow
% \baselineskip to stretch. Authors submitting work to the IEEE should note
% that IEEE rarely uses double column equations and that authors should try
% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
% packages (also by Sigitas Tolusis) as IEEE does not format its papers in
% such ways.
% Do not attempt to use stfloats with fixltx2e as they are incompatible.
% Instead, use Morten Hogholm'a dblfloatfix which combines the features
% of both fixltx2e and stfloats:
%
% \usepackage{dblfloatfix}
% The latest version can be found at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/dblfloatfix/

\usepackage{url}
\usepackage{glossaries}

\newacronym{APICSS}{APICSS}{Art Painting Image Color Semantics}
\newacronym{RGB}{RGB}{Red, Green, Blue}
\newacronym{HSL}{HSL}{Hue, Saturation, Luminance}
\newacronym{HSV}{HSV}{Hue, Saturation, Value}

% correct bad hyphenation here

\begin{document}

\title{Digital Analysis of Paintings}

\author{\IEEEauthorblockN{Alexander David Brown}
\IEEEauthorblockA{Computer Science,\\
Aberystwyth University,\\
Penglais,
Aberystwyth,\\
Ceredigion,\\
Wales SY23 3DB\\
Email: adb9@aber.ac.uk}
}

\maketitle

\begin{abstract}
%TODO Write this once the paper is complete
\end{abstract}

% no keywords

\IEEEpeerreviewmaketitle

\section{Introduction}
Digital image processing is a field which encourages the crossover of different
scientific disciplines, often biological research and computer vision align to
automate the collection and analysis of plant growth. However, art and
computing are fields which, at first glance, have very little in common.

But a deeper look unveils a plethora of different avenues from detecting
forgeries to being able to date an artists work within a catalogue of their
known pieces.

This survey paper will unearth some techniques which can be applied to the
digital analysis of paintings, as well as existing research which has already
applied computer vision to the field of art.

\section{Colour Analysis}
Digital images are typically considered to be a matrix of pixels, where each
pixel contains information about the colour of that place in the image.

Colours can be represented in numerous different ways, but are typically three
to four bytes; for example the \gls{RGB} colour space uses one byte for the
levels of red, one for green and one for blue. The fourth byte is unlikely to
be considered in artwork as it usually represents the transparency, known as
the alpha channel, of the pixel.

A single byte colour space purely focuses on the intensity of a pixel, in
actuality this represents a grayscale image.

Analysing colour is the base of all analysis techniques in image processing,
the value(s) of a pixel with regards to its location and neighbours can be used
to build up some very complex knowledge about the image. This section will deal
specifically with the colours which an artist uses within their work, such as
looking at the distribution of colours across a work, rather than using colour
information to determine textures.

\subsection{Colour Distribution}
% Paper 7

The distribution of colour across an image can give a surprising amount of
information about that image, especially given that colour is very subjective
to an individual.

Ivonna, Stanchev and Dimitrov investigated the colour
distribution within the works of over 100 artists for several different
countries and periods, using a system named \gls{APICSS}\cite{ivanova2008analysis}.

Existing systems already statistically analysed colours within an image and
some even used classifiers, typically a naive Bayesian classifier as it fits
well with statical methods, to perform extra analysis.

Unlike these existing system, \gls{APICSS} focused primarily on the \gls{HSL}
colour space as it is the closest representation to an artist's colour wheel.
\gls{HSV} was also considered, but because the lightness is not symmetrical
within the \gls{HSV} colour space, it is less useful for direct comparison.

As with any system involving classification, a feature space is needed to
represent an item within the data set. In the case of \gls{APICSS} the feature
space is three dimensional, one for each of the values in the colour space.

From this, a distance measure (in this case Euclidean distance) can be used to
compare different items.

\gls{APICSS} also take into account some of the metadata attached to each
painting, such as the movement and sub-movement during which the painting was
created. This allows for more statistical analyses to be performed on sub-sets
of the data set.

\gls{APICSS} appears to be a good system, which considers both metadata and
analysis gained from the digitisation of artwork. The paper considers a wide
range of artists and periods.

However, the use of a lossy image format (JPEG) could potentially skew the
results. The system itself is very basic, especially if it only considers
\gls{HSL} and metadata. This is further compounded by only considering thirteen
colours within the Hue (twelve separate colours and one achromatic).

Despite having a decent sized data set overall, some of the categories have
relatively small items within them.

Overall \gls{APICSS} is a good application of existing research, but uses very
basic analysis techniques on the images themselves.

\subsection{Pigment Mapping}
% Paper 3
Colour analysis can also give a glimpse into the colour pigments the artist
used to create the work. Using a database of known oil pigments and
\textit{a priori} knowledge, it is possible to work out which pigments are used
in a painting, and also to be able to show which parts of the painting are made
up from which pigment\cite{zhao2008investigation}.

This is done through a process known a multispectral imaging, the process of
capturing a different frequencies to provide a different spectrum of light and
enable more analysis.

This technique is especially powerful for segmenting different pigments in an
image as each pigment is made up of a different substance.

Because this process requires both hardware and software it make a large
different whether the materials a work is made up of are known. When known it
can be possible to find the combinations of these materials at a given point
and their concentrations.

When the materials are unknown, extra processing is required to decide which
materials the painting comprises of. This process uses spectral processing but
also requires prior knowledge of the artist and art tools.

In the experimental results this paper provides, the latter is achieved by a
study of another of Van Gogh's works from the same time.

Pigment mapping across every pixel is a costly operation in terms of memory and
CPU time. This research makes a reasonable assumption that a similar colour
will have been painted using similar pigments to reduce the amount of pigment
mapping which occurs. This does lead to unwanted side-effects, but the only way
to combat these is to increase the number of pixels which pixel-mapping.

To reduce the number of pixels considered for pigment mapping the image is
first partitioned into a number of smaller images to reduce processing time on
the next steps

Each of these partitioned regions is then split into a number of clusters based
on the colours of the pixels. This was done using supervised and unsupervised
algorithms. The unsupervised algorithm was very sensitive and had long running
times whilst the supervised algorithm requires a human observer.

Finally pixels within these clusters were then selected based on the similarity
of pixels within the cluster, the clusters and partitioned images can then be
stitched back together and rendered based on the pigment maps.

This research could be incredibly useful if it could be applied to the analysis
of artwork and the field of multispectral analysis has already had a lot of
success in other fields.

However, because this is both a hardware- and software-based system, it may not
be something which can be used with regularity as it requires the physical
presence of the paintings, which not all artwork analysis researchers have
access to.

It also requires a lot of a priori knowledge of the artist and calculating the
materials a painting consists of may be a difficult, if not impossible task,
depending on the artist.

In conclusion, this technique can provide a very powerful analysis of a
painting, but at the cost of requiring specialised hardware, the physical
painting and expert knowledge about the artist, so is a very manual process.

This technique could be improved by using a less naive clustering unsupervised
algorithm to remove some of the need for human intervention, but the need for
the physical painting will always be a limiting factor of this work.

\subsection{Difference Visualisation}
% Paper 6

\section{Texture Analysis}
Paintings are somewhat unlike the normal subject for image processing, whilst
most images are either simple two dimensional images, or two dimensional slices
of a three dimensional object. Paintings often thought of as two dimensional,
but with many paint types these paintings become three dimensional.

This aspect is lost in the digitisation of the painting; but there are still
ways of analysing the texture of the image.

\subsection{Steerable Filters}
% Paper 12

\subsection{Gabor Filters}
% Paper 13

\subsection{Histograms of Edge Orientated Gradients}
% Paper 8

\subsection{Multifractal Classification}
% Paper 5

\subsection{Texton-Based Analysis}
% Paper 11

\section{Statistical Analysis}

\subsection{Stylistic Analysis}
% Paper 9

\subsection{Authentication of Artwork}
% Paper 4

\subsection{Dating an Artist's Work}
% Paper AWESOME :D

\section{Brushstroke Analysis}

\subsection{Artistic Identification}
% Paper 1

\subsection{Rhythmic Brushstrokes}
% Paper 2

% conference papers do not normally have an appendix

% use section* for acknowledgement

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
\bibliography{bibliography}

% that's all folks
\end{document}